---
title: "Untitled"
author: "Olajuwon"
date: "10/23/2020"
output:
  word_document: default
  html_document: default
---


```{r message=FALSE}
library(tidyverse)
```


#Reads the data into R

```{r}
Lake<-read.csv("Lake.csv")

```



#I removed the first two columns with strings. this is to capture columns with numeric data

```{r}
Lake1<- data.frame(Lake[,3:43]) 
```


#This is to create a long data, it's easier to work with long data vs wide in certain scenarios.

Gather

```{r}
Lake_long<- gather(data= Lake, key= Variable, value= Value, -State, -MonitoringLocationName)


```

#Use mutate to do a log transformation of the Values to increase symmetry in the data

Transform

```{r}
Lake_long_log10<-mutate(Lake_long,Value= log10(Value+1))






```



```{r}
library(ggridges)
```


#Plot to visualize variables by state. 

```{r}
ggplot(Lake_long_log10)+
  geom_density_ridges(aes(y=State, x=Value, fill=State), alpha=0.5)+
  facet_wrap(~Variable, scales="free")+theme_classic()

ggsave("Lake_long.png", height=10, width=15, dpi=600)
```
#I prefer installing packages outside my code chunks. For other parts of the analysis we require devtools and ggbiplot
install.packages("devtools")
install_github("vqv/ggbiplot")

```{r}

library("devtools")
library("ggbiplot")
```


#Another set of data transformation while keeping our duplicate.






```{r}
Lake_log10<-Lake
Lake_log10[,3:43]<-data.frame(log10((Lake[,3:43])+1))




```


#TIme to reveal our components

```{r}
Lake_pca_log<-prcomp(Lake_log10[,3:43],scale.=TRUE)
```


Summary

```{r}
summary(Lake_pca_log)


```





#Some data visualization to see how our screeplot looks like

```{r}
plot(Lake_pca_log)
```



Perm
```{r}
pca_eigenperm<-function(data,  nperm=1000) {
  pca_out<-prcomp(data, scale.=T)
  eigenperm<-data.frame(matrix(NA, nperm, ncol(data)))
  n<-ncol(data)
  data_i<-data.frame(matrix(NA, nrow(data), ncol(data)))
  for (j in 1:nperm){
  for (i in 1:n) {
    data_i[,i]<-sample(data[,i], replace=F)
    }
  pca.perm<-prcomp(data_i, scale.=T)
  eigenperm[j,]<-pca.perm$sdev^2
  }
  
  colnames(eigenperm)<-colnames(pca_out$rotation)
  eigenperm
}
 
```
 
 #some permutations on our pca
 
```{r}
Lake_pca_perm<-pca_eigenperm(Lake_log10[,3:43], nperm=1000)

```
 
 

#99th percentile of each pc plus eigen values

```{r}

Lake_pca_rand99<-
  data.frame("Random_0.99"=sapply(Lake_pca_perm, quantile, 0.99)) %>% 
  mutate(PC=colnames(Lake_pca_log$rotation)) %>%  
  cbind(Eigenvalues=Lake_pca_log$sdev^2) 
  
 
```





#plot 

```{r}

Lake_pca_rand99_long<-
  gather(Lake_pca_rand99[1:9,], key=Variable, value=Value, -PC) 
```

#Plotting and saving our file using ggsave

```{r}
ggplot(Lake_pca_rand99_long, aes(PC, Value, fill=Variable))+
  geom_bar(stat="identity", position=position_dodge())+
  labs(y="Eigenvalue", x="", fill="")+
  theme_classic()

ggsave("fa_pca_rand99.png", height=5, width=7, dpi=600)
```

#time to view our PC Loadins


```{r}
Lake_loadings<-data.frame(Lake_pca_log$rotation[,1:5] )
Lake_loadings
```

```{r}
Lake_cor_loadings<- t(as.matrix(Lake_loadings)) * (Lake_pca_log$sdev)
Lake_cor_loadings<- t(Lake_cor_loadings)
Lake_cor_loadings
```






#Behold our BIplot that explains how various physicochemical factors are related in Texas and Oklahoma
```{r}

library(ggbiplot)

ggbiplot(Lake_pca_log, choices=c(1,2),  obs.scale = 1, varname.size=2,  var.scale = 1, ellipse=TRUE, color=Lake_log10$State, group=Lake_log10$State)+
  labs(color="State")+
  theme_classic()
 ggsave("Lake_pca_plot_view.png", height=10, width=15, dpi=600)
```



